---
title: What makes a Movie have High Audience Ratings and High Profit? What is the
  Relatinship between a Movie's High Audience Ratings and High Profit?
author: "Marie Hasegawa - mhasegawa7045@floridapoly.edu"
df_print: paged
output:
  html_document:
    df_print: paged
  word_document: default
pdf_document: default
---

# Appendix 
- Appendix
- Library of Packages
- Project Overview
- Exploratory Data Analysis
  - `Profitability` Column
  - `AudienceScore` Column
- Methods
  - `Profitability` vs. `AudienceScore` Grouped by `HighProfit` 
  - `Profitability` and `AudienceScore` K-Means Clustering
    - Elbow Method and Silhouette method to determine optimal clusters
    - K-Means Clustering of k=2,3,4
  - `Profitability` Column
    - Decision Tree `Profitability`
      - With Categorical Variables
      - Removed Categorical Variables
    - Bagging
    - Random Forest
    - Random Forest with Boosting
    - Model Comparison High `Profitability`
  - `AudienceScore` Column
    - Decision Tree `Profitability`
      - With Categorical Variables
      - Removed Categorical Variables
    - Bagging
    - Random Forest
    - Random Forest with Boosting
    - Model Comparison High AudienceScore
- Conclusion
  


# Library of Packages

|**Variables** |**Description**               |
|:-------------|:-----------------------------|
|`Movie`       |Title of movie                |
|`LeadStudio`  |Studio that released the movie|
|`RottenTomatoes` |Rotten Tomatoes rating (reviewers)|
|`AudienceScore`|Audience rating (via Rotten Tomatoes)|
|`Story` |General theme - one of 21 themes      |
|`Genre`|Type of Movie: Action, Adventure, Animation, Comedy, Drama, Fantasy, Horror, Romance, or Thriller |
|`TheatersOpenWeek`|Number of screens for opening weekend|
|`BOAverageOpenWeek`|	Average box office income per theater - opening weekend|
|`DomesticGross`|Gross income for domestic viewers (in millions) |
|`ForeignGross`|Gross income for foreign viewers (in millions)|
|`WorldGross`|Gross income for all viewers (in millions)|
|`Budget`  |Production budget (in millions)|
|`Profitability` |WorldGross/Budget    |
|`OpeningWeekend`   |Opening weekend gross (in millions)|


```{r}
library(tidyverse)
library(ISLR)
library(caret)
library(tidyverse)
#install.packages("skimr")
library(skimr) 
library(rpart.plot)
library(pROC)
library(ggplot2)
#install.packages("factoextra")
library(factoextra)
#install.packages("gridExtra")
library(gridExtra)


movie <- read_csv("https://raw.githubusercontent.com/reisanar/datasets/master/HollywoodMovies.csv")
```


# Project Overview

The box office industry is one of the biggest global markets in the entertainment industry. According to IMDbPro, the industry has a yearly average box office revenue of around 11 to 14 million dollars and a yearly total gross of around 9 to 12 billion dollars.
The project wants to determine what makes a Hollywood movie have a high `Profitability` and what makes it have a high `AudienceScore`, and how `Profitability` and `AudienceScore` correlate with each other. Movie studios would find this information resourceful since it could help them determine the best kind of movies that will lead to larger profits and more positive feedback on their movies for better publicity.

This will be done with EDA, K-means clustering, decision trees, bagging, boosting, and random forests that will help determine the pattern for high-earning and highly-rated movies. This project will be done in RStudio and RCloud with the dataset, `HollywoodMovies`, provided by Dr. Reinaldo Sanchez-Arias.

# Explatory Data Analysis


```{r}
movie<-movie %>%
  na.omit()
sample_n(movie, size=20)
```

```{r}
movie %>%
  filter(Profitability>=400)
```


The movie(`HollywoodMovies`)  dataset originally had many NA values in all columns, which needed to be removed. The code above is used to remove NA values with `na.omit()` since it needs to be removed in order to train the code and create decision and random trees.



## `Profitability` Column

```{r}
summary(movie$Profitability)
```
The variable, `Profitability`, in the dataset is the division between `WorldGross` and `Budget` in millions. The variables is used to determine if the movie made a huge profit beween  `WorldGross` and `Budget` in millions. The variable, `Profitability`, will be used in this section to see what factors affect `Profitability`of a movie and how highly-profited movies and lowly profited movies differ from each other.
The variable, `HighProfit`, will be mutated from the variable, `Profitability`, in order to differ between high-profited movies and low-profited movies. Based on the summarization shown above, the project will define the variable,`HighProfit`, to determine  a movie is highly-profitable if `Profitability` has a value of 400 or more, since the mean is 356.04 and the `third-quarter` is 397.08. The `HighProfit` variable can not be in the thousands like in the `max value` of 6694.40, since those values are considered an outlier becuase there are too few points to consider it.


```{r}
raw_A<-ggplot(data=movie) +
  geom_histogram(aes(x=Profitability)) +
  labs(title="Bar Chart of Profitability")
```

```{r}
rawB_Prof<-ggplot(data=movie) +
  geom_histogram(aes(x=Profitability)) +
  xlim(0,2000) +
  labs(title="Bar Chart of Profitability Removed Outliers")
```

```{r}
movieB<-movie %>%
  mutate(HighProfit=ifelse(movie$Profitability >= 400, "Yes", "No")) 

mBHP_F<-movieB %>%
  ggplot(aes(Profitability, colour = HighProfit)) +
  geom_freqpoly() +
  xlim(0,2000) +
  labs(title="Frequency Plot of HighProfit")

mBHP_D<-movieB %>%
  ggplot(aes(Profitability, colour = HighProfit)) +
  geom_bar() +
  geom_density(aes(y = ..count.., group=HighProfit),color="black") +
  xlim(0,2000) +
  labs(title="Bar Chart of HighProfit with Density Analysis")

grid.arrange(raw_A,rawB_Prof,mBHP_F, mBHP_D)  
grid.arrange(raw_A,rawB_Prof,mBHP_F, mBHP_D, ncol=1)  

#aes(group=HighProfit),method="lm",color="black", size=0.3
```
As you can see in the histogram, `Bar Chart of Profitibility`, `Profitability` accumulates mostly in the range of `~0 to ~2000 and has outliers in the range of `~2000 to 6500. Due to this, the outliers will be omitted for the EDA and in the `Bar Chart of Profitibility`, so it can be easier to analyze the relationship between high-profited movies and low-profited movies. This can be seen in the graph shown below with x (`Profitability`) limited between 0 to 2000. 

The `Frequency Plot of HighProfit` and the `Bar Chart of HighProfit with Density Analysis` shows that low-profited movies occur more frequently compared to high-profited movies. The reason might be that a movie's success is dependent on the movie's advertisement, availability globally, and its name/branding recognition.

For example, the movie, **Shrek the Third** is an extremely profitable movie of 499.35 but it is a sequel of the other Shrek film and is based on a very popular icon, Shrek, which could contribute a film's `Profitability` that is not shown on the graph.



## `AudienceScore` Column

```{r}
summary(movie$AudienceScore)
```
The variable, `AudienceScore`, in the dataset is a collection of the Rotten Tomatoes ratings from the audience. The variable `HighAud` is used to determine if the movie made a good impression on the Rotten Tomatoes audience. The variable, `AudienceScore`, will be used in this project to see what factors affect `AudienceScore` ratings of a movie and how highly-rated movies based on audience reviews differ from lowly-rated films.
The variable, `HighAud`, will be mutated from the variable, `AudienceScore`, in order to differentiate between highly-rated by-audience movies  and low-rated movies. Based on the summarization shown above, the project will define the variable, `HighAud`, to determine a movie is highly-rated by critics if `AudienceScore` has a value of 70 or more since the `third-quarter` is 73. 
```{r}
raw_B<-ggplot(data=movie) +
  geom_histogram(aes(x=AudienceScore)) +
  labs(title="Bar Chart of HighAud ")
```


```{r}
movieB<-movieB %>%
  mutate(HighAud=ifelse(movie$AudienceScore >= 70, "Yes", "No")) 

Freq_mB<-movieB %>%
  ggplot(aes(AudienceScore, colour = HighAud)) +
  geom_freqpoly() +
  xlim(0,100) + 
  labs(title="Frequency Plot of HighAud with Density Analysis")

Bar_mB<-
  movieB %>%
  ggplot(aes(AudienceScore, colour = HighAud)) +
  geom_bar() +
 geom_density(aes(y = ..count.., group=HighAud),color="black") +
  xlim(0,100)+
  labs(title="Bar Chart of HighAud with Density Analysis")
  
grid.arrange(raw_B,Freq_mB, Bar_mB, nrow = 2)  
grid.arrange(raw_B,Freq_mB, Bar_mB)  

```
There is barely any difference between low-rated movies and highly-rated movies in the `Bar Chart of HighAud with Density Analysis`. There is a ~5-count difference between the highly rated movies and lowly rated movies in the `Frequency Plot of HighAud`.
However, the density in the `Bar Chart of HighAud with Density Analysis` shows that low-rated movies and highly-rated movies peak at the same `count` level of 12.
The `Frequency Plot of HighAud` and the `Bar Chart of HighAud with Density Analysis` shows that low-audience rated movies occur more frequently compared to high-profited movies.


# Methods

## `Profitability` vs. `AudienceScore` Grouped by `HighProfit` 
The purpose of this section is to see the relationship between `Profitability` and `AudienceScore` by clustering the points with `HighProfit`.


```{r}
movieB

movieBclusHighProf<-movieB %>%
ggplot(aes(x = AudienceScore, y = Profitability, color = as.factor(HighProfit))) + 
  geom_point() +
  geom_smooth(aes(group=HighProfit),method="lm",color="black", size=0.3)+
  labs(title = "ScatterPlot of AudienceScore vs Profitability with Linear Method")

movieBclusHighProf_2000<-movieB %>%
ggplot(aes(x = AudienceScore, y = Profitability, color = as.factor(HighProfit))) + 
  geom_point() +
  geom_smooth(aes(group=HighProfit),method="lm",color="black", size=0.3)+
  ylim(0,2000)+
  labs(title = "ScatterPlot of AudienceScore vs Profitability with Linear Method without Outliers")
grid.arrange(movieBclusHighProf,movieBclusHighProf_2000, nrow = 1)
grid.arrange(movieBclusHighProf,movieBclusHighProf_2000)
```

This section shows that `AudienceScore` for highly-profited movies has correlated to a bigger increase in profit. This might have to do with audiences consisting of more people compared to critics and that more people come to the movie as an audience more often than as a critic. 

The `ggplots` shown above shows the relationship between `AudienceScore` and `Profitability` that is clustered with `HighProfit`. 

The  `ScatterPlot of AudienceScore vs Profitability with Linear Method` for lowly profited movies has a slightly, positive slope, while highly-profited movies have also a negative slope. This shows that as the `AudienceScore` for lowly profited increases the `Profitability` increases and for highly-profited increases then there is an increase in `Profitability`. 

It shows that the  `ScatterPlot of AudienceScore vs Profitability with Linear Method without Outliers` for lowly profited movies have a positive slope, while highly-profited movies have also a positive slope. This shows that as the `AudienceScore` for lowly profited increases the `Profitability` increases and for highly-profited increases then there is an increase in `Profitability`.  

These scatterplots show that as more highly-profiting outliers appear in the graph, the slope of lowly-profiting movies and the slope of highly-profiting movies decreases, which could mean that highly-profiting outlier has a negative correlation to the slope of lowly-profiting movies and the slope of highly-profiting movies.

## `Profitability` and `AudienceScore` K-Means Clustering 
### Elbow Method and Silhouette method to determinine optimal clusters
```{r}
myvars<-c("RottenTomatoes", "AudienceScore", "TheatersOpenWeek", "DomesticGross", "ForeignGross","WorldGross","Budget","Profitability","OpeningWeekend","BOAvgOpenWeekend")
```
The code above is used to remove the categorical variables, `Movie` and `LeadStudio`, so the dataset can be used for kmeans, since kmeans does not respond to datasets with columns being a collection of string values.

```{r}
movieB<-movieB[myvars] %>%
  na.omit()

movieB<-data.frame(movieB)

```

```{r}
set.seed(123)
# Elbow method
ELBmovie<-fviz_nbclust(movieB, kmeans, method = "wss") +
    geom_vline(xintercept = 4, linetype = 2)+
  labs(subtitle = "Elbow method")

# Silhouette method
SLmovie<-fviz_nbclust(movieB, kmeans, method = "silhouette")+
  labs(subtitle = "Silhouette method")
set.seed(123)
grid.arrange(ELBmovie, SLmovie, nrow = 2)

```
In order to find the most optimal number clusters to use for k-means clustering. The `elbow method` and the `silhouette method` will be used to determine the best number of clusters.
The `elbow method` computes k-means clustering for every different value of k. Then for every k value has the total within-cluster sum of squares (wss) calculated. Based on this, the best cluster for the `elbow method` would be 4.
The  `silhouette method` measures the quality of a cluster which determines how well each object lies within their cluster. Based on this and the `Optimal number of clusters Silhouette Method`, the most optimal number of clusters is 2. 
The project will choose the k values 2, 3, and 4 for K-means clustering since they are the most optimal k-values based on the `elbow method` and the `silhouette method`.

### K-Means Clustering of k=2,3,4
```{r}
myvars<-c("RottenTomatoes", "AudienceScore", "TheatersOpenWeek", "DomesticGross", "ForeignGross","WorldGross","Budget","Profitability","OpeningWeekend","BOAvgOpenWeekend")
```
The code above is used to remove the categorical variables, `Movie` and `LeadStudio`, so the dataset can be used for kmeans, since kmeans does not respond to datasets with columns being a collection of string values.

```{r}
movieB<-movieB[myvars] %>%
  na.omit()

movieB<-data.frame(movieB)

```


```{r}
movieB_clust<-movieB[,5]
mcls2<-kmeans(x=movieB_clust[!is.na(movieB_clust)], centers=2)
mcls3<-kmeans(x=movieB_clust[!is.na(movieB_clust)], centers=3)
mcls4<-kmeans(x=movieB_clust[!is.na(movieB_clust)], centers=4)
```


```{r}
mcls2
```
`mcls2` is a k-means clustering with 2 clusters or centers of size 58 and 533. The cluster mean of 1 is 471.00779 and the cluster mean of 2 is 57.84004. The sum of squares in cluster 1 is 3761951 and in cluster 2 is 1794220.

```{r}
mcls3
```
`mcls3` is a k-means clustering with 3 clusters or centers of size 84, 13, and 494. The cluster mean of 1 is 298.49229, the cluster mean of 2 is 795.62923, and the cluster mean 3 of 46.01348. The sum of squares in cluster 1 is 864482.1, cluster 2 is 1754172.2, and cluster 3 is 828400.7.

```{r}
mcls4
```

`mcls4` is a k-means clustering with 4 clusters or centers of size 115, 420, 45, and 11. The cluster mean of 1 is 153.82391,  cluster mean of 2 is 32.55724, cluster mean of 3 is 391.06960, and cluster mean of 4 is 835.02455. The sums of squares in cluster 1, 2, 3, and 4 are 268727.5,  285472.8,  292001.1, and 1641993.1.

```{r}
mB2<-fviz_cluster(mcls2, geom = "point", data=movieB)+ggtitle("Hollywood Movie Cluster Plot with k=2")
mB3<-fviz_cluster(mcls3, geom = "point", data=movieB)+ggtitle("Hollywood Movie Cluster Plot with k=3")
mB4<-fviz_cluster(mcls4, geom = "point", data=movieB)+ggtitle("Hollywood Movie Cluster Plot with k=4")

grid.arrange(mB2, mB3, mB4, nrow = 2)
grid.arrange(mB2, mB3, mB4, ncol = 1)

```
The `Hollywood Movie Cluster Plot with k=2` shows 2 clusters. 
Cluster 1 takes up the `Dim1 (53.1%)` range between -17 to -2 and `Dim2 (17.2%)`  range between -2.5 to 6.  Cluster 2 takes up the `Dim2 (53.1%)` range between -17 to -2 and `Dim1 (17.2%)`  range between -2.5 to 6.

The `Hollywood Movie Cluster Plot with k=3` shows 3 clusters. 
Cluster 1 has `Dim1 (53.1%)` range between -10 to 0 and `Dim2 (17.2%)`  range between 0 to 5. The Cluster 2 has `Dim1 (53.1%)` range between -0.5 to 0.5 and `Dim2 (17.2%)`  range between -2.5 to 3.75. Cluster 3 has a ` Dim1 (53.1%)` range between -0.5 to 0.5 and `Dim2 (17.2%)` range between -2.5 to 5.

The `Hollywood Movie Cluster Plot with k=4` shows 4 clusters. 
Cluster 1 has `Dim1 (53.1%)` range between -6 to 0 and `Dim2 (17.2%)`  range between -2.5 to 5. The Cluster 2 has `Dim1 (53.1%)` range between -2.5 to 5 and `Dim2 (17.2%)`  range between -2.5 to 4.75. Cluster 3 has a ` Dim1 (53.1%)` range between -10 to 7.5 and `Dim2 (17.2%)` range between -2.5 to 5. Cluster 4 has a ` Dim1 (53.1%)` range between -6 to 0 and `Dim2 (17.2%)` range between -2.5 to 5.

As you can see, the clusters in all k-means clustering are overlapped with the other clusters within their graphs. Based on this, `Hollywood Movie Cluster Plot with k=3` and `Hollywood Movie Cluster Plot with k=2` have better clusters than `Hollywood Movie Cluster Plot with k=4` since there is too much overlap to analyze. `Hollywood Movie Cluster Plot with k=3` would be the most optimal since it has more than 2 variables and has less overlap than `Hollywood Movie Cluster Plot with k=4`.




## `Profitability` Column
### Decision Tree `Profitability`
#### With Categorical Variables

This section and the next section are meant to be used as a comparison between the decision trees of `Profitability` that have categorical variables or not.  

```{r}
movie<- movie %>%
  na.omit()
movie
```

The `movie`(`HollywoodMovies` dataset) dataset originally had many NA values in all columns, which needed to be removed. The code above is used to remove NA values with `na.omit()` since it needs to be removed to train the code and create decision and random trees.


```{r}
HighProfit <- ifelse(movie$Profitability >= 400, "Yes", "No")

movie<-data.frame(movie, HighProfit)

movie$Profitability<-as.factor(movie$Profitability)
```

The code above creates a new variable called `HighProfit` that will be used to determine if a movie has a profit of more than 400 in the `Profitability` column. The `HighProfit` column will be determining this by labeling movies with string values `Yes` or  `No` depending on the movie's profitability.

```{r}
#remove.packages("Rcpp")
#install.packages('Rcpp')
library(Rcpp)
tree.movieProf<-train(HighProfit ~ . -Profitability, data=movie, method="rpart")
tree.movieProf
```
The variable, `HighProfit` was trained with the dataset `movie` (the stored`Hollywood` dataset) and without the variable, `Profitability`. As you can see that there are relatively low `Kappa` scores that suggest that the two raters do have a weak agreement with the two raters when using nominal scores. 
The most optimal model has an `Accuracy` of 0.8425314, `cp` of 0.08163265, and a `Kappa` of 0.5445023, which is not great because of `Accuracy` being average and the low `Kappa`. This training method used 25 Bootstrapped reps for resampling.  

```{r}
rpart.plot(tree.movieProf$finalModel)
```
For this section, the categorical variable was not removed so that the decision trees with categorical variables and without categorical variables can be compared.

The decision tree with categorical variables has that the root node shows that 25% of movies have high `Profitability` of  `more than 400. 

> When `OpenProfit` is less than 63, then there 11% of movies have high `Profitability` of more than 400, and 76% of them have an `OpenProit` of less than 63. 

>If movies have `OpenProfit` of more than 63 and a Foreign Gross of more than 25 million dollars, then 90% of movies have a `Profitability` of more than 400, and 15% of those movies have a `ForeignGross` of more than 25 million.  

>However, if movies have `OpenProfit` of more than 63 and a Foreign Gross of less than 25 million dollars, then 29% of movies have a `Profitability` of more than 400, and 9% of those movies have a `ForeignGross` of more than 25 million.  

This decision tree is somewhat poor since the depth of the decision tree is too small and too few variables to consider. That is why the next section is the same as this section but the decision tree does not include the categorical variables.

These rules can be shown at the bottom.

```{r}
rpart.rules(tree.movieProf$finalModel)
```



#### Removed Categorical Variables
This section and the previous section are meant to be used as a comparison between the decision trees of `Profitability` that have categorical variables or not. This section removes the categorical variables, `Movie` and `LeadStudio`.

```{r}

movie <- read_csv("https://raw.githubusercontent.com/reisanar/datasets/master/HollywoodMovies.csv")
myvars<-c("RottenTomatoes", "AudienceScore", "TheatersOpenWeek", "DomesticGross", "ForeignGross","WorldGross","Budget","Profitability","OpeningWeekend","BOAvgOpenWeekend")
```

```{r}
movie<-movie[myvars] %>%
  na.omit()

movie<-data.frame(movie)

movie<-movie %>%
  mutate(HighProfit=ifelse(movie$Profitability >= 400, "Yes", "No")) %>%
  dplyr::select(-Profitability)

movie<-movie %>%
  mutate(HighProfit=as.factor(HighProfit))
```

The `movie` dataset originally had many NA values in all columns, which need to be removed. The code above is used to remove NA values with `na.omit()` since it needs to be removed to train the code and create decision and random trees. The code above also creates a new variable called `HighProfit` that will be used to determine if a movie has a profit of more than 400 in the `Profitability` column. The `HighProfit` column will be determining this by labeling movies with string values of `Yes` or  `No` depending on the movie's profitability.


```{r}
inTrain<-createDataPartition(y=movie$HighProfit, p=0.75, list=FALSE)
train <- movie[inTrain,]
test <- movie[-inTrain,]
```

```{r}
set.seed(217)
cvcontrol <- trainControl(method = "repeatedcv", number = 10, allowParallel = TRUE)

```
The two chunks above create training and testing sets for the `HighProfit` variable in the dataset, `movie`. The split is needed to train the data in order to better create decision trees. A `trainControl` variable was made to be used on the `train` variable shown at the bottom. The variable, `cvcontril` is made with the `trainControl()` method that uses the method `repeatedcv`, which is `cv` that repeats, 10-folds classification, and allows the code to run parallel.

```{r}
dt_movie<- train(HighProfit ~ . , data = train, method = "rpart", 
                     trControl = cvcontrol, tuneLength = 10)
dt_movie

```
The `train()` method-used model uses a 10-fold classification done at 1 time and includes 2 classes that define if the movie is of `HighProfit` or not.
The most optimal model, according to the `CART` table shown above, has a `cp` value of 0.008823529, `accuracy` of 0.8852791, and a `Kappa` of 0.6925392, which is average and not the absolute best in determining the best decision process to get the most `profitable` movie.

```{r}
plot(dt_movie)
```

The plot above shows the relationship between the `complexity parameter` and the `Accuracy (Repeated Cross-Validation)`. The graph shows it has a negative relationship. This suggests that as the `complexity parameter` increases, the `accuracy` decreases.

```{r}
rpart.plot(dt_movie$finalModel)

```
The decision tree shown above does not include categorical variables that would hinder the creation of a precise decision tree. This is a much superior decision tree compared to the previous section since it has more depth than the decision tree with categorical variables.

The best chance to have movies with `Profitability` of more than 400 million dollars `to occur 100%` happens in 1 scenario:

> If `DomesticGross` has an average of fewer than 134 million dollars, `Budget` is more than 12 million dollars and less than 18 million dollars, and `WorldGross` is more than 65 million dollars and less than 113 million dollars, then 100% of movies are highly-profitable and 2% of them has a `WorldGross` greater than 65 million dollars. 


```{r}
tree_classTrain<-predict(dt_movie, type="raw")
head(tree_classTrain)
```

```{r}
confusionMatrix(train$HighProfit, tree_classTrain)
```

The accuracy of the `confusion matrix and Statistics` is relatively great since it is 0.941  correct.  The `95% confident` that the interval (0.9196, 0.9581). The `kappa` is relatively good and has a value of 0.8504, which means that there is somewhat near-perfect agreement. This confusion matrix also claims that the p-value is less than 2e-16  which is very low and that this model does greatly go against the null hypothesis. It also has no positive class.


```{r}
tree_classTest<-predict(dt_movie, newdata=test, type="prob")
tree_classTest
```

```{r}
ROC_curve_dt<-roc(test$HighProfit, tree_classTest[,"Yes"])
plot(ROC_curve_dt)
auc(ROC_curve_dt)
```
The graph shown above shows that the Area under the curve takes up 0.8701 which means that the performance of the model can fairly distinguish between the positive and negative classes. This graph will be saved in order to compare how well all the decision tree techniques interpreted the test data. 
The slope at specificity 1 with a sensitivity of 0 to 0.8 is undefined since the graph in this area is vertical, but has a positive slope after the specifity of 1 to 0. 

### Bagging

```{r}
bag_movie<-train(HighProfit~.,data=train, method="treebag", trControl=cvcontrol, importance=TRUE)

bag_movie
```
The `bagged CART` model has two predictors that determine if the movie has a high profit of more than 400 million dollars.
The `bagging` will be used to improve the accuracy of the algorithms used in the machine learning algorithm. The `bag_movie` contains 2 classes called `Yes` and `No` that indicate if the movie agrees with `HighProfit`. It also has only one optimal model with an `accuracy` of 0.9153866 and a `kappa` value of 0.7750693

```{r}
plot(varImp(bag_movie))
```

The `varImp `plot shown above shows that `WorldGross`, `DomesticGross`, `Budget`, and `ForeignGross` are the most important to the bagged `HighProfit` variable with importance values of 100, 95, 64, and 60. This means that the project should consider these variables as important to the training and testing of the variable `HighProfit`.

```{r}
bagg_pred<-predict(bag_movie, newdata = test, type="raw")
head(bagg_pred)
```

```{r}
confusionMatrix(test$HighProfit, bagg_pred)
```
The `accuracy` of the `confusion matrix and Statistics` is relatively great since it is 0.9183 correct.  The `95% confident` that the interval is (0.9014, 0.9698). The `kappa` is relatively average and has a value of 0.7863, which means that there is some perfect agreement. This confusion matrix also claims that the p-value is less than 9.709e-10 which is very low and that this model does greatly go against the null hypothesis. This `confusion matrix and Statistics` is great because of the high `Accuracy` and high `Kappa`.


```{r}
bagg_probs <- predict(bag_movie, newdata = test, type = "prob")
head(bagg_probs)
```

```{r}
ROC_curve_bagg <- roc(test$HighProfit, bagg_probs[ , "Yes"])
plot(ROC_curve_bagg)
```

```{r}
auc(ROC_curve_bagg)
```

The graph shown above shows that the Area under the curve takes up 97.77% which means that the performance of the model can easily distinguish between the positive and negative classes. This graph will be saved in order to compare how well all the decision tree techniques interpreted the test data. The slope at specificity 1 with a sensitivity of 0 to 0.8 is undefined since the graph in this area is vertical, but has a zero slope after the specificity of 1 to 0 because that section is flat. 

### Random Forest 
```{r}
rf_HighProfit<-train(HighProfit~., data=train, method="rf", trControl=cvcontrol, importance =TRUE)
rf_HighProfit
```
The training method for the `random forest` has 2 classes that determine if `HighProfit` is high by 400 million dollars or more. This training method was also done in a 10-fold classification dine 1 time
The `random forest` method will be used to improve the accuracy of the algorithms used in the machine learning algorithm. It contains 2 classes called `Yes` and `No` that indicate if the movie agrees with `HighProfit`. It also has an optimal model with an `accuracy` of 0.9283410,  a `kappa` value of 0.8118733, and a `mtry` of 9. This `random forest` is relatively slightly above average since the `Accuracy` is high and the `Kappa` is average.  

```{r}
rf_pred<-predict(rf_HighProfit, newdata=test, type="raw")
confusionMatrix(test$HighProfit, rf_pred)
```
The `accuracy` of the `confusion matrix and Statistics` is relatively great since it is 91.35% correct.  The `95% confident` that the interval (0.8667, 0.9479). The kappa is relatively average and has a value of 0.7724, which means that there is some agreement. This `confusion matrix` also claims that the p-value is 8.096e-09  which is very low and that this model does greatly go against the null hypothesis. The `confusion matrix and Statistics` is relatively great due to the high `Accuracy` and the high `Kappa`. 

```{r}
rf_probs<-predict(rf_HighProfit, newdata = test, type = "prob")
head(rf_probs)
```

```{r}
ROC_curve_rf<-roc(test$HighProfit, rf_probs[,"Yes"])
plot(ROC_curve_rf)
auc(ROC_curve_rf)
```
The graph shown above shows that the Area under the curve takes up 98.09% which means that the performance of the model can easily distinguish between the positive and negative classes which are almost perfect. This graph will be saved in order to compare how well all the decision tree techniques interpreted the test data. The AUC of `ROC_curve_rf` is relatively higher compared to the past 2 methods which means that this is more useful than the other two methods since its overall performance would be better.


### Random Forest with Boosting
```{r}
gbm_movie<-train(HighProfit~., data=train,
                 method="gbm", verbose=F, trControl=cvcontrol)
gbm_movie
```
The Stochastic Gradient Boosting with 10-fold cross-validation shows the most optimal model has 150 trees, an `interactive depth` of 2, shrinkage of 0.1, `Accuracy` of 0.9298771, and `Kappa` of 0.8137917, which can be said as a relatively good model because of its `Kappa` and `Accuracy`. This is done to the variable, `HighProfit`, since the project needs to determine what makes a movie be highly-profitable. 

```{r}
gbm_preds <- predict(gbm_movie, newdata = test, type = "raw")
# get confusion matrix
confusionMatrix(test$HighProfit, gbm_preds)
```
The `confusionMatrix and Statistics` for this section shows that the `accuracy` of 0.9327, the  `95% confidence interval` of (0.8896, 0.9627), `Kappa` value of 0.8209, and the p-value of 2.591e-10, which means that it goes against the null hypothesis.

```{r}
gbm_probs<-predict(gbm_movie, newdata = test, type = "prob")
ROC_curve_gbm<-roc(test$HighProfit, gbm_probs[,"Yes"])
plot(ROC_curve_gbm)
auc(ROC_curve_gbm)
```

The graph shown above shows that the Area under the curve takes up 97.8%` which means that the performance of the model can easily distinguish between the positive and negative classes and has a high performance. This graph will be saved in order to compare how well all the decision tree techniques interpreted the test data. The AUC of `ROC_curve_rf` is relatively higher compared to the past 3 methods which means that this is more useful than the other two methods since its overall performance would be better.


### Model Comparison High-`Profitability`
```{r}
# using the `add = TRUE` option for each plot after the first one:
plot(ROC_curve_dt, col = "black", legacy.axes = T, )  # black
plot(ROC_curve_bagg, add = TRUE, col = "red") # color red for bagging
plot(ROC_curve_rf, add = TRUE, col = "green") # color green is for random forest
plot(ROC_curve_gbm,add = TRUE, col = "orange") # color orange is for boosting

auc(ROC_curve_dt)
auc(ROC_curve_bagg)
auc(ROC_curve_rf)
auc(ROC_curve_gbm)
```

The graphs shown above show the `AUC-ROC` of the `decision tree`, `bagging`, `rainforest`, and `rain forest`. 
The best method for the variable, `HighProfit`, is `rain forest` which is shown in the green curve, since it has the highest AUC-ROC by a small margin, which we will consider highly. The worst method for the variable, `HighProfit` is `decision tree` since it has the lowest AUC-ROC by a large margin, which we will not consider. All methods gave an indefinite slope for a certain section and suddenly have a positive slope.


## `AudienceScore` Column
### Decision Tree `AudienceScore`
#### With Categorical Variables
This section and the next section are meant to be used as a comparison between the decision trees of `AudienceScore` that have categorical variables or not. 

```{r}
movie <- read_csv("https://raw.githubusercontent.com/reisanar/datasets/master/HollywoodMovies.csv")

movie<- movie %>%
  na.omit()
movie
```
The `movie`(`HollywoodMovies` dataset) dataset originally had many NA values in all columns, which needed to be removed. The code above is used to remove NA values with `na.omit()` since it needs to be removed in order to train the code and create decision and random trees.

```{r}
HighAud <- ifelse(movie$AudienceScore >= 70, "Yes", "No")

movie<-data.frame(movie, HighAud)

movie$AudienceScore<-as.factor(movie$AudienceScore)
```

The code above creates a new variable called `HighAud` that will be used to determine if a movie has an audience rating of more than 70 in the `AudienceScore` column. The `HighAud` column will be determining this by labeling movies with string values `Yes` or  `No` depending on the movie's audience score.

```{r}
tree.movieAud<-train(HighAud ~ . -AudienceScore, data=movie, method="rpart")
tree.movieAud
```
The `CART training method` model uses 25 bootstraped reps for resampling. The variable, `HighAud` was trained with the dataset `movie` (the stored`Hollywood` dataset) and without the variable, `AudienceScore`. As you can see that there are relatively low `Kappa` scores that suggest that the two raters do have a weak agreement with the two raters when using nominal scores. 
The most optimal model has an `Accuracy` of 0.7935805, `cp` of 0.03125, and a `Kappa` of 0.5064600, which is not a very good `CART` model to use.



```{r}

rpart.plot(tree.movieAud$finalModel)
```
For this section, the categorical variable was not removed so the decision trees with categorical variables and without categorical variables can be compared with this section and the next section.

The decision tree with categorical variables has that the root node shows that 32% of movies have a high `AudienceScore` of more than 70. 

>When `RottenTomatoes` is less than 65, then 15% of movies have a high `AudienceScore` of more than 70, and 67% of them have a `RottenTomatoes` of less than 65. 

>If movies have a `RottenTomatoes` value of more than 81, then 87% of movies have an `AudienceScore` of more than 70, and 17% of those movies have a `RottenTomatoes` of more than 81.  

>If movies have a `RottenTomatoes` value of more than 65 but less than 81 and `WorldGross` of less than 215 million dollars, then 35% of movies have a `WorldCross` of more than 70, and 12% of those movies have a `WorldClass` of more than 215.  

>If movies have a `RottenTomatoes` value of more than 65 but less than 81 and a `WorldGross` of more than 215 million dollars, then 89% of movies have a `WorldCross` of more than 70, and 5% of those movies have a `WorldClass` of more than 215. 

These rules can be shown at the bottom.

```{r}
rpart.rules(tree.movieAud$finalModel)
```



#### Removed Categorical Variables

This section and the previous section are meant to be used as a comparison between the decision trees of `AudienceScore` that have categorical variables or not. This section removes the categorical variables, `Movie` and `LeadStudio`.

```{r}
movie <- read_csv("https://raw.githubusercontent.com/reisanar/datasets/master/HollywoodMovies.csv")
myvars<-c("RottenTomatoes", "AudienceScore", "TheatersOpenWeek", "DomesticGross", "ForeignGross","WorldGross","Budget","Profitability","OpeningWeekend","BOAvgOpenWeekend")
```


```{r}
movie<-movie[myvars] %>%
  na.omit()

movie<-data.frame(movie)

movie<-movie %>%
  mutate(HighAud=ifelse(movie$AudienceScore >= 70, "Yes", "No")) %>%
  dplyr::select(-AudienceScore)

movie<-movie %>%
  mutate(HighAud=as.factor(HighAud))
```

The `movie` dataset originally had many NA values in all columns, which need to be removed. The code above is used to remove NA values with `na.omit()` since it needs to be removed to train the code and create a decision and random trees. The code above also creates a new variable called `HighAud` that will be used to determine if a movie has an audience score of `more than 70 in the `AudienceScore` column. The `HighAud` column will be determining this by labeling movies with string values of `Yes` or  `No` depending on the movie's audience reviews.

```{r}
inTrain<-createDataPartition(y=movie$HighAud, p=0.75, list=FALSE)
train <- movie[inTrain,]
test <- movie[-inTrain,]
```

```{r}
set.seed(217)
cvcontrol <- trainControl(method = "repeatedcv", number = 10, allowParallel = TRUE)

```
The two chunks above create the training and testing sets for the `HighAud` variable in the dateset, `movie`. The split is needed to train the data to better create decision trees. A `trainControl` variable was made to be used on the `train` variable shown at the bottom.

```{r}
dt_movie<- train(HighAud ~ . , data = train, method = "rpart", 
                     trControl = cvcontrol, tuneLength = 10)
dt_movie
```
The `CART model` uses the `HighAud` variable and is resampled with 10-fold classification and cross-validation at one time. 
The most optimal model, according to the `CART` table shown above, has a `cp` value of 0, `accuracy` of 0.8150674, and a `Kappa` of 0.57963801, which is not great since the `Accuracy` is average and the `Kappa` is really low.

```{r}
plot(dt_movie)
```

The plot above shows the relationship between the `complexity parameter` and the `Accuracy (Repeated Cross-Validation)`.The graph's slope becomes 0 and the line becomes flat between the `complexity parameters` of 0.05 to 0.4 with an `Accuracy` of ~0.80. However, `Accuracy` immediately drops by 0.11 units between the `complexity parameter` of 0.4 to 0.43.
This suggests that the `Accuracy` normalizes in a certain `complexity parameter` frame, but suddenly the `Accuracy` of the model suddenly drops. 

```{r}
rpart.plot(dt_movie$finalModel)
```
The decision tree shown above does not include categorical variables that would hinder the creation of a precise decision tree. This is a much superior decision tree compared to the previous section since it has more depth than the decision tree with categorical variables.

The most optimal models have an 89% chance are highly rated movies:
> If `RottenTomatoes` has an average of less than 69 but more than 42, `Profitability` of more than 280 million dollars, and `DomesticGross` is more than 117 million dollars, then 89% of movies have high-profitability and 4% of those movies have a `profitability` of more than 280 million dollars. 

> If `RottenTomatoes` has an average of more than 82, then 89% of movies have high-profitability and 17% of those movies have a `RottenTomatoes` of more than 82. 

> If `RottenTomatoes` has an average of more than 69 but less than 82 and `WorldGross` of more than 173 million dollars, then 89% of movies have high-profitability and 6% of those movies have a `WorldGross` of more than 173 million dollars. 

```{r}
tree_classTrain<-predict(dt_movie, type="raw")
head(tree_classTrain)
```

```{r}
confusionMatrix(train$HighAud, tree_classTrain)
```
The `confusionMatrix and Statistics` for this section shows that the `accuracy` of 0.8804, the  `95% confidence interval` of (0.8524, 0.9047), `Kappa` value of 0.7313, and the p-value of less than 2e-16, which means that it goes against the null hypothesis. This model is not the best since its `Kappa` value indicates the overall agreement level is average.  


```{r}
tree_classTest<-predict(dt_movie, newdata=test, type="prob")
tree_classTest
```

```{r}

ROC_curve_dt<-roc(test$HighAud, tree_classTest[,"Yes"])
plot(ROC_curve_dt)
auc(ROC_curve_dt)
```

The graph shown above shows that the Area under the curve takes up 78.47% which means that the performance of the model has an average time distinguishing between the positive and negative classes. Overall, this model would be an okay graph to use. This graph will be saved in order to compare how well all the decision tree techniques interpreted the test data. The reason it has such a low AUC is that the area between the black and gray graphs is smaller compared to the previous AUC graphs.

### Bagging

```{r}
bag_movie<-train(HighAud~.,data=train, method="treebag", trControl=cvcontrol, importance=TRUE)

bag_movie
```
The `bagging` will be used to improve the accuracy in the algorithms used of the machine learning algorithm. The `bag_movie` contains 2 classes called `Yes` and `No` that indicate if the movie agrees with `HighAud`. It also has an optimal model with an `accuracy` of 0.8308756 and a `kappa` value of 0.6164977, which is inadequate since the `Accuracy` and the `Kappa` are so low, which says that the overall agreement level is poor.

```{r}
plot(varImp(bag_movie))
```

The `varImp `plot shown above shows that `RottenTomatoes`, `DomesticGross`, and `WorldGross` are the most important to the bagged `HighAud` variable with importance values of 100, 80, and 60. This means that the project should consider these variables as important to the training and testing of the variable `HighAud`.

```{r}
bagg_pred<-predict(bag_movie, newdata = test, type="raw")
head(bagg_pred)
```

```{r}
confusionMatrix(test$HighAud, bagg_pred)
```
The `accuracy` of the `confusion matrix and Statistics` is relatively average since it is 0.8606 accurate. The `95% confidence interval` is (0.8059, 0.9046). The `kappa` is relatively low and has a value of 0.6756, which means that there is a poor overall agreement level. This confusion matrix also claims that the p-value is 1.308e-06 which is very low and that this model does greatly go against the null hypothesis. However, this model is not a good choice since it has a low `Accuracy` and low `Kappa`.


```{r}
bagg_probs <- predict(bag_movie, newdata = test, type = "prob")
head(bagg_probs)
```

```{r}
ROC_curve_bagg <- roc(test$HighAud, bagg_probs[ , "Yes"])
plot(ROC_curve_bagg)
```

```{r}
auc(ROC_curve_bagg)
```
The graph shown above shows that the Area under the curve takes up 87.37% which means that the performance of the model can distinguish between the positive and negative classes. This graph will be saved in order to compare how well all the decision tree techniques interpreted the test data. This `AUC model` is better than the previous `AUC model` because it has a larger area between the black and gray graphs which means that it has a better overall performance.

### Random Forest 
```{r}
rf_HighAud<-train(HighAud~., data=train, method="rf", trControl=cvcontrol, importance =TRUE)
rf_HighAud
```
The `random forest` method will be used to improve the accuracy of the algorithms used in the machine learning algorithm. It contains 2 classes called `Yes` and `No` that indicate if the movie agrees with `HighAud`, and used a resampling of 10-fold classification that is processed once. 
It also has an optimal model with an `accuracy` of 0.8405530, `kappa` value of 0.6375480, and a `mtry` of 5. This `Random Forest` is relatively average due to the average `Accuracy` and low `Kappa` value.


```{r}
rf_pred<-predict(rf_HighAud, newdata=test, type="raw")
confusionMatrix(test$HighAud, rf_pred)
```
The `accuracy` of the `confusion matrix and Statistics` is relatively average since it is 0.8558  accurate.  The `95% confidence interval` is (0.8005, 0.9005). The `kappa` is relatively average and has a value of 0.6656, which means that there is a poor average overall agreement level. This confusion matrix also claims that the p-value is 8.616e-05 which is very low and that this model does greatly go against the null hypothesis. However, this model is not a good choice since it has a low `Accuracy` and low `Kappa`.



```{r}
rf_probs<-predict(rf_HighAud, newdata = test, type = "prob")
head(rf_probs)
```

```{r}
ROC_curve_rf<-roc(test$HighAud, rf_probs[,"Yes"])
plot(ROC_curve_rf)
```

```{r}
auc(ROC_curve_rf)
```

The graph shown above shows that the Area under the curve takes up 89.67% which means that the performance of the model can easily distinguish between the positive and negative classes. This graph will be saved in order to compare how well all the decision tree techniques interpreted the test data.

### Random Forest with Boosting
```{r}
gbm_movie<-train(HighAud~., data=train,
                 method="gbm", verbose=F, trControl=cvcontrol)
gbm_movie
```
The Stochastic Gradient Boosting with resampling of 10-fold cross-validation shows the most optimal model has 100 trees, an `interactive depth` of 1, shrinkage of 0.1, `Accuracy` of 0.8439228, and `Kappa` of 0.6440923, which can be said as a relatively average model because of its low `Kappa` and average `Accuracy`.



```{r}
gbm_preds <- predict(gbm_movie, newdata = test, type = "raw")
# get confusion matrix
confusionMatrix(test$HighAud, gbm_preds)
```
The `confusionMatrix and Statistics` for this section show that the `Accuracy` of `85.1%`, the  `95% confidence interval` of (0.7952, 0.8964), the `Kappa` value of 0.6581, and the p-value of 4.388e-07, which means that it goes against the null hypothesis.

```{r}
gbm_probs<-predict(gbm_movie, newdata = test, type = "prob")
ROC_curve_gbm<-roc(test$HighAud, gbm_probs[,"Yes"])
plot(ROC_curve_gbm)
```

```{r}

auc(ROC_curve_gbm)
```

The graph shown above shows that the Area under the curve takes up 0.9035 which means that the performance of the model can easily distinguish between the positive and negative classes. This graph will be saved in order to compare how well all the decision tree techniques interpreted the test data. This `AUC model` is better than the previous `AUC models` because it has a larger area between the black and gray graphs which means that it has a better overall performance.

### Model Comparison High `AudienceScore`
```{r}
# using the `add = TRUE` option for each plot after the first one:
plot(ROC_curve_dt, col = "black", legacy.axes = T, )  # black
plot(ROC_curve_bagg, add = TRUE, col = "red") # color red for bagging
plot(ROC_curve_rf, add = TRUE, col = "green") # color green is for random forest
plot(ROC_curve_gbm,add = TRUE, col = "orange") # color orange is for boosting
auc(ROC_curve_dt)
auc(ROC_curve_bagg)
auc(ROC_curve_rf)
auc(ROC_curve_gbm)
```

The graphs shown above show the `AUC-ROC` of the `decision tree`, `bagging`, `random forest`, and `random forest with boosting`. The best method for the variable, `HighAud`, is `random forest with Boostingt` which is shown on the orange graph since it has the highest AUC-ROC by a small margin.


# Conclusion

One of the biggest global markets in the entertainment industry is the box office industry. According to `IMDbPro`, the industry has a yearly average box office revenue of around 11 to 14 million dollars and a yearly total gross of around 9 to 12 billion dollars. The project wanted to determine what makes a Hollywood movie have a high `Profitability` what makes it have a high `AudienceScore`, and how `Profitability` and `AudienceScore` correlate with each other. This was done with EDA, clustering, decision trees, bagging, boosting, and random forests that will help determine the pattern for high-earning and highly-rated movies.

The `Profitability vs. AudienceScore Grouped by HighProfit`section shows a model of the relationship between `Profitability` and `AudienceScore` that is clustered with the variable, `HighProfit`, which determines if a movie is highly-profitable or not. The `ScatterPlot of AudienceScore vs Profitability with Linear Method` for lowly profited movies has a slightly, positive slope, while highly-profited movies have also a negative slope. The `ScatterPlot of AudienceScore vs Profitability with Linear Method without Outliers` for lowly profited movies have a positive slope, while highly-profited movies have also a positive slope. These scatterplots show that as more highly-profiting outliers appear in the graph, the slope of lowly-profiting movies and the slope of highly-profiting movies decreases, which could mean that highly-profiting outlier has a negative correlation to the slope of lowly-profiting movies and the slope of highly-profiting movies.

The `Profitability and AudienceScore K-Means Clustering` uses the elbow method and K-means clustering for `AudienceScore`. The project chose the k values 2, 3, and 4 for K-means clustering since they are the most optimal k-values based on the elbow method and the silhouette method. The clusters in all k-means clustering are overlapped with the other clusters within their graphs. Based on this, `Hollywood Movie Cluster Plot with k=3` and `Hollywood Movie Cluster Plot with k=2` have better clusters than `Hollywood Movie Cluster Plot with k=4` since there is too much overlap to analyze. `Hollywood Movie Cluster Plot with k=3` would be the most optimal since it has more than 2 variables and has less overlap than `Hollywood Movie Cluster Plot with k=4`.

The model comparison of the `Profitibility`column includes decision trees (with and without categorical variables), bagging, random forest, and random forest with boosting. The `AUC-ROC Curve of Random Forest with Boosting` shows that the best method for the variable, `HighAud`, is a random forest with boosting since it has the highest AUC-ROC by a small margin.

The model comparison of the `AudienceScore`column includes decision trees (with and without categorical variables), bagging, random forest, and random forest with boosting. The `AUC-ROC Curve of Random Forest with Boosting` shows the AUC-ROC of the decision tree, bagging, random forest, and random forest. The best method for the variable, `HighAud`, is the random forest with boosting since it has the highest AUC-ROC by a small margin.

This project taught me that usually low `Kappa` has an average or low `Accuracy`. It also taught me that the `rpart()` and `rpart.plot()` makes unique decision tree in every re-run, which was difficult for me to deal with because I had to constantly change the text to correlate to the decision trees.

Movie studios would find this information useful since it could help them determine the best kind of movies that will lead to larger profits and more positive feedback on their movies and firms for better publicity and recognition.


# Resources

Fonseca, L. (2019, August 15). Clustering analysis in R using K-means. Retrieved April 24, 2021, from https://towardsdatascience.com/clustering-analysis-in-r-using-k-means-73eca4fb7967

Hidayatuloh, A. (n.d.). Visualize Clustering Using ggplot2. Retrieved April 24, 2021, from https://rpubs.com/aephidayatuloh/clustervisual

Michel Kana, P. (2020, February 24). Data clustering tutorial for advanced. Retrieved April 24, 2021, from https://towardsdatascience.com/clustering-for-data-nerds-ebbfb7ed4090

K-means cluster analysis. (n.d.). Retrieved April 24, 2021, from https://afit-r.github.io/kmeans_clustering

Rashmi, Kassambara 06 May 2020 The demo data used in this tutorial is available in the default installation of R. Juste type data(USArrests) Reply, &amp; Kassambara. (2018, October 21). K-Means clustering in R: Algorithm and practical examples. Retrieved April 24, 2021, from https://www.datanovia.com/en/lessons/k-means-clustering-in-r-algorith-and-practical-examples/

